---
title: "Inroduction to fars package"
author: "Faizan Khan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Introduction
This vignette shows how to use the `fars` package. The `fars` package has functions to read the data downloaded from the [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars), summarize the data, and map the locations of accidents.

## Installation
You can install this package directly from GitHub.

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("faizan-khan-iit/fars")
```

## Downloading Data

Go to the [FARS website](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars) and download the data in the current working directory. A sample file has been included in the package as an example in `inst/extdata`.

```{r, eval = FALSE}
file_2015 <- system.file("extdata", "accident_2015.csv.bz2", package = "fars")
```


## Reading the data
To read the downloaded data files, use the `fars_read_years()` function. Multiple years' data can also be read at once by passing a vector containing multiple values.
```{r "fars_read_years()", eval=FALSE}
# Read data for year 2015
fars_read_years(c(2015))

# Read data for years 2010 to 2014
fars_read_years(c(2010:2014))
```

We can see the accident cases for every month for the years provided.


## Summarized data
Use the `fars_summarize_years()` to read and summarize the data for the years provided.

```{r "fars_summarize_years()", eval=FALSE}
# Summarize data for year 2015
fars_summarize_years(c(2015))
```

The function returns a `tibble` with the number of cases per month for a given year.

## Plot the locations of accident on a map

Use `fars_map_state()` to plot the locations for the accidents for a particular state, defined by `state.num`. The `state.num` values are according to the FARS website.
```{r "fars_map_state()", eval=FALSE}
# Plot accident locations for 2015 for state number 15, Hawaii
fars_map_state(state.num = 15, year = 2015)
```
